---
title: "buoy_weight"
author: "Hayden Vega"
date: "2024-10-29"
output: html_document
description: An rmd with buoyant weight analysis of corals collected for the fish-regen project. buoy_data is raw buoyant weight data, and coral_delta is the data for final-initial values for variables buoyant_weight_g, dry_mass, and volume. 
---
### Setup
```{r}
library("tidyverse")
library("readxl")
library("janitor")
library('readxl')
library("here")
library("ggplot2")
```

```{r}
buoy_data <- read.csv(here("data", "fish_regen_buoyantweight (1).csv"))
coral_delta <- read.csv(here("data", "coral_delta.csv"))
```

```{r}
buoy_data%>% 
  mutate(density_stopper = (air_weight_g * 0.9965)/(air_weight_g - fresh_weight_g))%>%
  mutate(density_sw = (air_weight_g - salt_weight_g)/ (air_weight_g / density_stopper))%>%
  mutate(dry_mass = bouyantweight_g / (1 - (density_sw/density_aragonite)))%>%
  mutate(coral_vol = (dry_mass / (density_aragonite - density_sw))) -> skeletal.mass 
```
# Allometric linear models

This section contains code for allometric linear models. 


```{r}
     #                        Data Wrangling                       ~~~
     #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

#Filter data to exclude erroneous outlier (coral 45) and extranneous columns. Pivot to create columns for initial and final buoyant weight. 
log_data <- buoy_data %>%
  filter(coral_id != 45) %>%   #remove 45
  select(coral_id, date, bouyantweight_g, wound, fish) %>%  # keep only relevant columns
  pivot_wider(names_from = date, values_from = bouyantweight_g) %>% #initial and final to columns, values from buoyantweight
  rename(
    initial_weight = initial, #renaming for clarity
    final_weight = final
  ) %>% 
  mutate(delta_mass = final_weight - initial_weight) #calculating delta

# Filter data for unwounded corals and no fish treatment 
log_df <- log_data %>%
  #filter(wound == "No Wound", fish == "No Fish") %>%
  mutate(
    log_i_weight = log(initial_weight),
    log_f_weight = log(final_weight)
  )
#...............................................................................
#                                                                              .
#  Should we log only the "no wound, no fish" corals?                           .
#                                                                              .
#...............................................................................

b_values_df <- log_df %>%
  #group_by(coral_id) %>% #fish treatment specific
  do({
    model <- lm(log_f_weight ~ log_i_weight, data = .)
    tidy_model <- broom::tidy(model, conf.int = TRUE)
    intercept_row <- tidy_model[1, ]
    slope_row <- tidy_model[2, ]
    tibble(
      intercept = intercept_row$estimate,
      intercept_conf.low = intercept_row$conf.low,
      intercept_conf.high = intercept_row$conf.high,
      b_value = slope_row$estimate,
      b_conf.low = slope_row$conf.low,
      b_conf.high = slope_row$conf.high
    )
  })

log_df <- log_df %>%
  cross_join(b_values_df)

log_df %>%
  ggplot(aes(x = log_i_weight, y = log_f_weight)) +
  geom_point(alpha = 0.7, color = "steelblue") +
  geom_abline(aes(intercept = intercept, slope = b_value), color = "darkred", size = 1) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "grey50") +
  labs(
    title = "Allometric Scaling",
    x = "Log Initial Weight",
    y = "Log Final Weight"
  ) +
  theme_bw(base_size = 14)

```

## Linear model of growth ratio as a function of wound and fish treatments

```{r}
#Note, no tank random effect because it's not in the data. RUH ROH
log_df <- log_df %>%
  mutate(growth_ratio = final_weight / initial_weight)

growth_lm_interaction <- lm(growth_ratio~fish * wound , data = log_df )
summary(growth_lm_interaction)
anova(growth_lm_interaction, test = "Chisq")


growth_lm_no_int <- lm(growth_ratio~ fish + wound , data = log_df)
anova(growth_lm_no_int, test = "Chisq")


growth_no_wound <- lm(growth_ratio ~ fish , data = log_df)
anova(growth_no_wound, test = "Chisq")

growth_no_fish <- lm(growth_ratio~wound , data = log_df)
anova(growth_no_fish, test = "Chisq")

#...............................................................................
#                                                                              .
#  It seems that neither fish nor wounding has a significant effect on growth  .
#                                                                              .
#...............................................................................


```
### Quantifying the best fitting model formula

This section contains my poor attempt to try and use the framework in simulating_wound_data.qmd to view how well the differing formulas fit the trends.
My graph currently yields a big old blank. 
```{r}
lrt_int <- anova(growth_lm_no_int, growth_lm_interaction, test = "Chisq")

# B) Test for fish main effect: does adding fish improve fit?
lrt_f   <- anova(growth_no_fish, growth_lm_no_int, test = "Chisq")

# C) Test for wound_size_cat main effect: does adding wound_size_cat improve fit?
lrt_s   <- anova(growth_no_wound, growth_lm_no_int, test = "Chisq")


coef_df <- tidy(growth_lm_interaction, effects = "fixed") %>%
  rename(
    term     = term,
    Estimate = estimate,
    SE       = std.error
  )

lrt_stats <- tribble(
  ~test_term,          ~LR.chisq,       ~Df, ~p_LRT,
  "Interaction",        lrt_int$Chisq[2], lrt_int$Df[2], lrt_int$`Pr(>Chisq)`[2],
  "Fish effect",        lrt_f$Chisq[2],   lrt_f$Df[2],   lrt_f$`Pr(>Chisq)`[2],
  "Wound size category",lrt_s$Chisq[2],   lrt_s$Df[2],   lrt_s$`Pr(>Chisq)`[2]
)

# 5c. Map each coefficient term to its corresponding LRT test label:
#     - fishpresent → "Fish effect"
#     - wound_size_catlarge → "Wound size category"
#     - fishpresent:wound_size_catlarge → "Interaction"
results_df <- coef_df %>%
  mutate(
    test_term = case_when(
      term == "fishNo Fish"                            ~ "Fish effect",
      term == "woundSmall"                    ~ "Wound size category",
      term == "fishNo Fish:woundSmall"        ~ "Interaction",
      TRUE                                             ~ NA_character_
    )
  ) %>%
  # Join with LRT stats by matching on test_term
  left_join(lrt_stats, by = "test_term") %>%
  # We only keep rows where a test_term exists (drop the intercept)
  filter(!is.na(test_term)) %>%
  # Select and order the columns we want in the final table
  select(term, Estimate, SE, LR.chisq, Df, p_LRT)

results_df %>%
  rename(
    ChiSq = LR.chisq,
    p     = p_LRT
  ) %>%
  kable(
    caption = "Table 1. GLMM Fixed Effects and Manual LRT χ² Tests",
    digits  = c(NA, 2, 2, 2, 0, 3)
  ) %>%
  kable_styling(full_width = FALSE, position = "center")

#  select(species, summary) %>%
 # unnest(summary)
```


```{r}


log_data <- buoy_data %>%
  filter(coral_id != 45) %>%
  pivot_wider(
    id_cols = coral_id,
    names_from = date,
    values_from = bouyantweight_g
  )

# Filter data for unwounded corals and no fish treatment
allometry_df <- df %>%
  filter(wound_raw == "not wounded", treatment == "no fish") %>%
  mutate(
    log_i_weight = log(i.weight),
    log_f_weight = log(f.weight)
  )
log_data %>% 
  ggplot(aes(initial, final))+
  geom_point()

0.991493 + .002019 * 2


#This graph 
skeletal.mass %>% 
  ggplot(aes(date, bouyantweight_g, color = fish, fill = wound))+
  geom_boxplot()
```


#graph change of drymass and volume
```{r fig.width = 5, fig.height = 5}

coral_delta %>% 
  ggplot(aes(fish, dry_mass_delta, color = wound))+
  geom_boxplot()+
  coord_cartesian(ylim = c(0, 2))

```

```{r}
coral_delta %>% 
  ggplot(aes(fish, vol_delta, color = wound))+
  geom_boxplot()+
  coord_cartesian(ylim = c(0, 2))
```

