---
title: "pam_code"
author: "Hayden Vega & Adrian Stier"
date: "2025-05-30"
output: html_document
description: An analysis of PAM data from the fish regen project. Included are visualizations, linear models, ANOVAs, Tukey tests, and randomization tests. 
---

# Setup
```{r}
#............................packages............................
library(janitor)
library(tidyverse)
library(viridis)
library(lme4)
library(ggeffects)
library(sjlabelled)
library(DHARMa)
library(parameters)
library(effects)
library(MuMIn)
library(here)
library(tidyverse)
library(gt)
library(broom.mixed)
library(ggpubr)


#..........................loading data..........................
pam_healed <- read.csv(here("data", "pam_healed.csv"))

pam_undisturbed_healed <- read.csv(here("data", "pam_undisturbed_healed.csv"))

#.......................loading aesthetics.......................
wound_palette <- c("Large" = "#ffbf00", 
                   "Small" = "#be8333")
fish_palette <- c("Fish" = "#2b4b52", 
                  "No Fish" = "#ad301a")

```


## Visualizing Fv/Fm by wound and fish presence

### Fv_Fm as a function of fish presence--colored by wound
```{r}
ggplot(pam_healed, aes(x = fish, y = fv_fm, fill = wound)) +
  geom_boxplot() +
  #geom_smooth(method = "lm", se = FALSE, linewidth = 1.1) +
  scale_fill_manual(values = wound_palette) +
  labs(
    title = "Coral Photosynthetic Efficiency as a Function of Fish Presence",
    subtitle = "Colored by Wound Treatment",
    x = "Fish Presence",
    y = "Photosynthetic Efficiency (fv/fm)",
    color = "Wound Treatment"
  ) +
  theme_pubr(base_size = 14)+
  theme(
    plot.title = element_text(hjust = 1), 
    plot.subtitle = element_text(hjust= -0.05)
  )

ggsave(
  here("figures", "photoeff_vs_fish.png"),
  width = 8,
  height = 6,
  dpi = 300
)
```

### Fv_Fm as a function of wound presence--colored by fish presence
```{r}
ggplot(pam_healed, aes(x = wound, y = fv_fm, fill = fish)) +
  geom_boxplot() +
  #geom_smooth(method = "lm", se = FALSE, linewidth = 1.1) +
  scale_fill_manual(values = fish_palette) +
  labs(
    title = "Coral Photosynthetic Efficiency as a Function of Wound Size",
    subtitle = "Colored by Fish Presence",
    x = "Wound Size",
    y = "Photosynthetic Efficiency (fv/fm)",
    color = "Fish Presence"
  ) +
  theme_pubr(base_size = 14)

ggsave(
  here("figures", "photoeff_vs_wound.png"),
  width = 8,
  height = 6,
  dpi = 300
)

#Based on both graphs above, I think there may be a significant interaction between large wounds and fish presence.


# Enforce correct factor ordering for all downstream plots
pam_healed <- pam_healed %>%
  mutate(
    wound = factor(wound, levels = c("Small", "Large")),
    fish = factor(fish, levels = c("No Fish", "Fish"))
  )
ggplot(pam_healed, aes(x = wound, y = fv_fm, fill = fish)) +
  geom_violin(position = position_dodge(0.9), alpha = 0.3, color = NA) +
  geom_boxplot(
    position = position_dodge(0.9), 
    width = 0.3, 
    outlier.shape = NA, 
    color = "black",
    alpha = 0.8
  ) +
  geom_jitter(
    aes(color = fish),
    position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.9),
    size = 2, 
    alpha = 0.5
  ) +
  scale_fill_manual(values = fish_palette) +
  scale_color_manual(values = fish_palette) +
  labs(
    title = "Coral Photosynthetic Efficiency vs Wound Size",
    subtitle = "Colored by Fish Presence",
    x = "Wound Size",
    y = "Photosynthetic Efficiency (Fv/Fm)",
    fill = "Fish Presence",
    color = "Fish Presence"
  ) +
  theme_bw(base_size = 18) +
  theme(
    plot.title = element_text(size = 20, face = "bold"),
    plot.subtitle = element_text(size = 16),
    axis.text = element_text(size = 16),
    axis.title = element_text(size = 18),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 15)
  )

ggsave(
  here("figures", "photoeff_vs_wound_violin_screen.png"),
  width = 7, height = 5, dpi = 150
)

```

## Linear Models and Anova for Fv/Fm
### Fitting Mixed Effect Models
```{r}

# Full Model OH YEAH
mod_int <- lmer(fv_fm ~ fish * wound + (1 | tank), data = pam_healed)

summary(mod_int)

# Additive model without interaction
mod_add <- lmer(fv_fm ~ fish + wound + (1 | tank), data = pam_healed)

# Fish-only model
mod_fish <- lmer(fv_fm ~ fish + (1 | tank), data = pam_healed)

# Wound-only model
mod_wound <- lmer(fv_fm ~ wound + (1 | tank), data = pam_healed)

# Null model (intercept + tank)
mod_null <- lmer(fv_fm ~ 1 + (1 | tank), data = pam_healed)

```

### Likelihood tests
```{r}
lrt_interaction <- anova(mod_add, mod_int, test = "Chisq")
lrt_fish <- anova(mod_wound, mod_add, test = "Chisq")
lrt_wound <- anova(mod_fish, mod_add, test = "Chisq")
```

### Summarize results in a table
```{r}
lrt_table <- tibble(
  Test = c("Interaction (Fish × Wound)", "Fish Effect", "Wound Effect"),
  ChiSq = c(
    lrt_interaction$Chisq[2],
    lrt_fish$Chisq[2],
    lrt_wound$Chisq[2]
  ),
  Df = c(
    lrt_interaction$Df[2],
    lrt_fish$Df[2],
    lrt_wound$Df[2]
  ),
  p_value = c(
    lrt_interaction$`Pr(>Chisq)`[2],
    lrt_fish$`Pr(>Chisq)`[2],
    lrt_wound$`Pr(>Chisq)`[2]
  )
) %>%
  mutate(across(where(is.numeric), round, 3))

# Format with gt
lrt_table %>%
  gt() %>%
  tab_header(
    title = "Table #. Likelihood Ratio Tests for Effects on fv/fm",
    #subtitle = paste("Scaled growth:", expression((final - initial)/initial^b))
  ) %>%
  cols_label(
    Test = "Model Comparison",
    ChiSq = "Chi-squared",
    Df = "Degrees of Freedom",
    p_value = "P-value"
  ) %>%
  fmt_number(columns = vars(ChiSq, p_value), decimals = 3) %>%
  tab_options(
    table.font.size = 14,
    heading.title.font.size = 16,
    heading.subtitle.font.size = 13
  )


```

### Averaging values 

```{r}
pam_avg_wound <- pam_healed %>%
  group_by(wound) %>%
  summarize(mean_fvfm = mean(fv_fm, na.rm = TRUE), 
  sd_fvfm = sd(fv_fm, na.rm = TRUE),
            n = n()) 
pam_avg_fish <- pam_healed %>% 
  group_by(fish) %>% 
  summarize(mean_fish = mean(fv_fm, na.rm = TRUE), 
  sd_fish = sd(fv_fm, na.rm = TRUE), 
            n = n())

pam_avg <- pam_healed %>% 
  group_by(fish, wound) %>% 
  summarize(mean_fv_fm = mean(fv_fm, na.rm = TRUE), 
            sd = sd(fv_fm, na.rm = TRUE), 
            n = n())
```
#Models with FV/fM and continuous wound size (initial_wound_size)

```{r}
rate_full <- lmer(
  fv_fm ~ fish * initial_wound_size + (1 | tank),
  data = pam_healed
)


rate_no_interaction <- lmer(
  fv_fm ~ fish + initial_wound_size + (1 | tank),  # 
  data = pam_healed
)


rate_no_fish <- lmer(
  fv_fm ~ initial_wound_size + (1 | tank),  # 
  data = pam_healed
)

rate_no_wound <- lmer(
  fv_fm ~ fish + (1 | tank),  
  data = pam_healed
)


lrt_interaction_zero_rate <- anova(rate_no_interaction, rate_full, test = "Chisq")
lrt_fish_zero_rate <- anova(rate_no_fish, rate_no_interaction, test = "Chisq")
lrt_wound_zero_rate <- anova(rate_no_wound, rate_no_interaction, test = "Chisq")

lrt_table <- tibble(
  Test = c("Interaction (Fish × Wound)", "Fish Effect", "Wound Effect"),
  ChiSq = c(lrt_interaction_zero_rate$Chisq[2],
            lrt_fish_zero_rate$Chisq[2],
            lrt_wound_zero_rate$Chisq[2]),
  Df = c(lrt_interaction_zero_rate$Df[2],
         lrt_fish_zero_rate$Df[2],
         lrt_wound_zero_rate$Df[2]),
  p_value = c(lrt_interaction_zero_rate$`Pr(>Chisq)`[2],
              lrt_fish_zero_rate$`Pr(>Chisq)`[2],
              lrt_wound_zero_rate$`Pr(>Chisq)`[2])
) %>%
  mutate(across(where(is.numeric), round, 3))

lrt_table %>%
  gt() %>%
  tab_header(
    title = "Table #. Likelihood Ratio Tests for Effects on Healing Rate",
  ) %>%
  cols_label(
    Test = "Model Comparison",
    ChiSq = "Chi-squared",
    Df = "Degrees of Freedom",
    p_value = "P-value"
  ) %>%
  fmt_number(columns = vars(ChiSq, p_value), decimals = 3) %>%
  tab_options(
    table.font.size = 14,
    heading.title.font.size = 16,
    heading.subtitle.font.size = 13
  )
```

# Old Code



### Boxplot of Pam for undisturbed tissue
```{r}


pam_undisturbed_healed %>% 
  ggplot(aes (x = fish, y = fv_fm, fill = fish)) +
  geom_boxplot()+
  theme_bw()

  ## Boxplot of PAM including x = wound and  fish = fill
pam_undisturbed_healed %>% 
  ggplot (aes (x = wound, y = fv_fm, fill = fish))+
  geom_boxplot()+
  theme_bw()

  ## Boxplot of PAM including x = fish and fill = wound

pam_undisturbed_healed %>% 
  ggplot (aes (x = fish, y = fv_fm, fill = wound))+
  geom_boxplot()+
  theme_bw()

pam_undisturbed_healed %>% 
  ggplot (aes (x = wound, y = fv_fm, fill = wound))+
  geom_boxplot()+
  theme_bw()

```


### Boxplot of PAM for healed tissue
```{r}
pam_healed %>% 
  ggplot(aes (x = fish, y = fv_fm, fill = fish)) +
  geom_boxplot()+
  theme_bw()

pam_healed %>% 
  ggplot(aes (x = fish, y = fv_fm, fill = wound)) +
  geom_boxplot()+
  theme_bw()

pam_healed %>% 
  ggplot(aes (x = wound, y = fv_fm, fill = wound)) +
  geom_boxplot()+
  theme_bw()

pam_healed %>% 
  ggplot(aes (x = wound, y = fv_fm, fill = fish)) +
  geom_boxplot()+
  theme_bw()






```


### Multiple Regression

```{r}
pam_all_regression <- lmer(fv_fm ~ wound + fish + (1 | f0), data = pam_undisturbed_healed)
summary(pam_all_regression)
anova(pam_all_regression)

pam_healed_regression <- lmer(fv_fm ~ wound + fish + (1 | f0), data = pam_healed)
summary(pam_healed_regression)
anova(pam_healed_regression)
```

### Anova
```{r}
pam_all_anova <- aov(fv_fm ~ wound * fish, data = pam_undisturbed_healed)

summary(pam_all_anova)

pam_healed_anova <- aov(fv_fm ~ wound * fish, data = pam_healed)
```
### Tukey
```{r}
pam_all_tukey <- TukeyHSD(pam_all_anova)
pam_all_tukey

pam_healed_tukey <- TukeyHSD(pam_healed_anova)
pam_healed_tukey
```


### R2 values
```{r}

MuMIn::r.squaredGLMM(pam_all_regression)

MuMIn::r.squaredGLMM(pam_healed_regression)
```

```{r}
pam_all_predictions <- ggpredict(pam_all_regression, terms = c("wound", "fish")) %>% 
  as_tibble() %>% 
  rename(wound = group)
```

### Visualize pam_all
```{r}
pam_all_plot <- ggplot() +
  
  # plot the raw data as points
  geom_point(data = pam_undisturbed_healed,
             aes(x = wound,
                 y = fv_fm),
             alpha = 0.5,
             color = "black",
             shape = 21) +
  
  # plot a 1:1 reference line
 # geom_abline(slope = 1, 
  #            intercept = 0,
   #           linetype = 2,
    #          linewidth = 1,
     #         color = "black") +
  
  # plot the confidence interval 
 # geom_ribbon(data = pam_all_predictions,
  #            aes(x = x,
   #               y = predicted,
    #              ymin = conf.low,
     #             ymax = conf.high),
      #        alpha = 0.2,
       #       fill = "black") +
  
  # plot the prediction line
#  geom_line(data = pam_all_predictions,
 #           aes(x = x,
  #              y = predicted),
   #         color = "black",
    #        linewidth = 1) +
  
  # labels
  labs(x = "Wound Intensity",
       y = "Photosynthetic Efficiency") 

pam_all_plot

```

### Plot TUKEY

```{r  fig.width = 20, fig.height = 40}
tukey_plot_all <- aov(fv_fm ~ wound:fish, data = pam_undisturbed_healed)

tukey_plot_all_plot <- TukeyHSD(tukey_plot_all)
plot(tukey_plot_all_plot)


tukey_plot_healed <- aov(fv_fm~wound:fish, data = pam_healed)

tukey_healed <- TukeyHSD(tukey_plot_healed)
plot(tukey_healed)

#Seems that large: no fish is significantly different from small and large with fish; suggests that as wound size increases (high metabolic stress) fish effects emerge. within treatments with fish did not vary. and treatments within small wounds did not vary significantly from each other. #supports positive gradient hypothesis #IS THIS SUPPORTED BY BUOYANT WEIGHT DATA?
```
### RANDOMIZATION TEST ARGHHHH
```{r}
#PAM first
#looking at the mean between wounds



#Actually randomize
null_photo_dist <- replicate(1000, {
  photo_means <- pam_healed %>% 
    mutate(wound = sample(wound, n())) %>% #this shuffles the selection of data
  group_by(wound) %>% 
  summarize(mean = mean(fv_fm))

 diff_means <- photo_means$mean[2]-
  photo_means$mean[1]
 
 diff_means
})

ggplot(tibble(null_photo_dist), aes(null_photo_dist)) +
  geom_histogram(bins = 20, 
                 color = "cornflowerblue", 
                 fill = NA)+
  geom_vline(xintercept = diff_means, 
             color = "firebrick")


```
### Means between fish within wounds
```{r}
large <- pam_healed %>% 
  filter(wound == "Large")

photo_means <- large %>% 
  group_by(fish) %>% 
  summarize(mean = mean(fv_fm))

#Actually randomize
nulll_photo_dist <- replicate(1000, {
  photo_means <- large %>% 
    mutate(fish = sample(fish, n())) %>% #this shuffles the selection of data
  group_by(fish) %>% 
  summarize(mean = mean(fv_fm))

 diff_means <- photo_means$mean[2]-photo_means$mean[1]
 
 diff_means
})

ggplot(tibble(nulll_photo_dist), aes(nulll_photo_dist)) +
  geom_histogram(bins = 20, 
                 color = "cornflowerblue", 
                 fill = NA)+
  geom_vline(xintercept = diff_means, 
             color = "firebrick")




sum(abs(diff_means) > abs(diff_means))/
  length(null_photo_dist)
```
### do it for small wounds
```{r}
small <- pam_healed %>% 
  filter(wound == "Small")

photo_means <- small %>% 
  group_by(fish) %>% 
  summarize(mean = mean(fv_fm))

#Actually randomize
nullll_photo_dist <- replicate(1000, {
  photo_means <- small %>% 
    mutate(fish = sample(fish, n())) %>% #this shuffles the selection of data
  group_by(fish) %>% 
  summarize(mean = mean(fv_fm))

 diff_means <- photo_means$mean[2] - photo_means$mean[1]
 
 diff_means
})

ggplot(tibble(nullll_photo_dist), aes(nullll_photo_dist)) +
  geom_histogram(bins = 20, 
                 color = "cornflowerblue", 
                 fill = NA)+
  geom_vline(xintercept = diff_means, 
             color = "firebrick")

```



### Do it for growth
```{r}
buoy_data <- read.csv("fish_regen_buoyantweight (1).csv")
```

```{r}
null_growth_dist <- replicate(1000, {
  photo_means <- small %>% 
    mutate(fish = sample(fish, n())) %>% #this shuffles the selection of data
  group_by(fish) %>% 
  summarize(mean = mean(fv_fm))

 diff_means <- photo_means$mean[2] - photo_means$mean[1]
 
 diff_means
})

ggplot(tibble(nullll_photo_dist), aes(nullll_photo_dist)) +
  geom_histogram(bins = 20, 
                 color = "cornflowerblue", 
                 fill = NA)+
  geom_vline(xintercept = diff_means, 
             color = "firebrick")
```

### something New
```{r}
null_photo_dist <- replicate(1000, {
  photo_means <- pam_healed %>%
    mutate(wound = sample(wound, n()), #this shuffles the selection of data
                     fish_presence = sample(fish, n())) %>% #this shuffles the selection of for fish presence
  group_by(wound) %>%
  summarize(mean = mean(fv_fm))

 diff_means <- photo_means$mean[2]-
  photo_means$mean[1]

 diff_means
})

ggplot(tibble(null_photo_dist), aes(null_photo_dist)) +
  geom_histogram(bins = 20,
                 color = "cornflowerblue",
                 fill = NA)+
  geom_vline(xintercept = diff_means,
             color = "firebrick")
```