---
title: "wound_closure_analysis"
output: html_document
---


# Setup
```{r}
#............................packages............................
library(janitor)
library(tidyverse)
library(viridis)
library(lme4)
library(ggeffects)
library(sjlabelled)
library(DHARMa)
library(parameters)
library(effects)
library(MuMIn)
library(here)
library(tidyverse)
library(gt)
library(broom.mixed)
library(ggpubr)
library(sysfonts)
library(showtext)
library(waffle)
library(glmmTMB)
#..............................data..............................
wound_closure <- read.csv(here("data", "fish_regen_mastersheet_wound closure_necrosis_sa - mastersheet.csv"))

percent_closure <- read.csv(here("data", "final_wound_size_dascyllus_project - Sheet1.csv"))
  

#.........................data wrangling.........................
wound_closure_clean <- wound_closure %>% 
   filter(!is.na(wound_close), wound_close != "") %>% 
  mutate(wound_close = ifelse(wound_close == "no-barely" | wound_close == "no - barely", "no", wound_close)) %>% 
  mutate(fish = ifelse(fish == 1, "Fish", "No Fish")) %>% 
  mutate(wound_close = ifelse(wound_close == "yes", "Yes", "No")) %>% 
  mutate(wound = ifelse(wound == 1, "Small", "Large")) %>% 
  mutate(wound_close = as.character(wound_close))

wound_closure_glm <- wound_closure %>% 
  filter(!is.na(wound_close), wound_close != "") %>% 
  mutate(wound_binary = ifelse(wound_close == "yes", 1, 0))

wound_prop <- wound_closure_clean %>% 
  group_by(wound, fish) |>
  summarize(prop_healed = mean(wound_close == "Yes"))

percent_closure <- percent_closure %>% 
    filter(percent.healed >= 0) %>% 
  mutate(fish = as.numeric(fish)) %>% 
  mutate(fish = ifelse(fish == 1, "Fish", "No Fish")) %>% 
  mutate(wound = ifelse(wound == 1, "Small", "Large")) 

percent_closure$tank <- as.factor(percent_closure$tank)

percent_closure$prop <- percent_closure$percent.healed/100 

percent_closure$prop_wound_remaining <- 1 - percent_closure$prop

#...........................aesthetics...........................
wound_palette <- c("Small" = "#ffbf00", 
                   "Large" = "#be8333")
fish_palette <- c("Fish" = "#2b4b52", 
                  "No Fish" = "#ad301a")
font_add_google(name = "Josefin Sans", family = "josefin")
showtext_auto()
theme_set(theme_bw() +
            theme(panel.grid = element_blank(), 
                  plot.background = element_blank()))
```
# Initial Plotting

## Wound Prop by Fish Presence -- colored by wound
```{r}

wound_prop %>% 
  ggplot(aes(fish, prop_healed, fill = wound))+
  geom_col(position = "dodge")+
  labs(x = "Fish Presence", 
       y = "Proportion Healed", 
       fill = "Wound Size",
       title = "Proportion of Healed Wounds as a Function of Fish Presence",
       subtitle = "Colored by Wound")+
  scale_fill_manual(values = wound_palette)+
  theme(
    legend.title = element_text(family = "josefin"), 
    axis.title.x = element_text(family = "josefin"), 
    axis.title.y = element_text(family = "josefin"), 
    plot.title = element_text(family = "josefin"), 
    plot.subtitle = element_text(family = "josefin"))

ggsave(
  here("figures", "healed_prop_vs_fish.png"),
  width = 8,
  height = 6,
  dpi = 300
)
```


## Prop final wound size by fish and wound
```{r}
percent_closure %>% 
  ggplot(aes(fish, prop_wound_remaining, fill = wound))+
  geom_boxplot()+
  labs(x = "Fish Presence", 
       y = "Percent Healed", 
       fill = "Wound Size",
       title = "Final Percent Healed as a Function of Fish Presence",
       subtitle = "Colored by Wound")+
  scale_fill_manual(values = wound_palette)+
  theme(
    legend.title = element_text(family = "josefin"), 
    axis.title.x = element_text(family = "josefin"), 
    axis.title.y = element_text(family = "josefin"), 
    plot.title = element_text(family = "josefin"), 
    plot.subtitle = element_text(family = "josefin"))


```

## Wound Prop by Wound Size -- colored by fish presence 
```{r}
wound_prop %>% 
  ggplot(aes(wound, prop_healed, fill = fish))+
  geom_col(position = "dodge")+
  labs(x = "Wound Size", 
       y = "Proportion Healed", 
       fill = "Fish Presence",
       title = "Proportion of Healed Wounds as a Function of Wound Size",
       subtitle = "Colored by Fish Presence")+
  scale_fill_manual(values = fish_palette)+
  theme(
    legend.title = element_text(family = "josefin"), 
    axis.title.x = element_text(family = "josefin"), 
    axis.title.y = element_text(family = "josefin"), 
    plot.title = element_text(family = "josefin"), 
    plot.subtitle = element_text(family = "josefin"))

ggsave(
  here("figures","healed_prop_vs_wound.png"),
  width = 8,
  height = 6,
  dpi = 300
)

# Load libraries
library(tidyverse)
library(here)

# --------------------------------------------
# 1. Prep waffle grid
# --------------------------------------------

wound_counts_long <- wound_closure_clean %>%
  mutate(outcome = ifelse(wound_close == "Yes", "Healed", "Not Healed")) %>%
  group_by(wound, fish, outcome) %>%
  summarise(n = n(), .groups = "drop")

waffle_grid <- wound_counts_long %>%
  uncount(n)

waffle_grid <- waffle_grid %>%
  group_by(wound, fish, outcome) %>%
  mutate(id = row_number()) %>%
  group_by(wound, fish) %>%
  mutate(square_y = row_number()) %>%
  ungroup()

waffle_grid <- waffle_grid %>%
  mutate(group_id = paste(fish, wound, sep = " - ")) %>%
  mutate(group_id = factor(group_id, 
                            levels = c("Fish - Large", "No Fish - Large", "Fish - Small", "No Fish - Small"))) 

# --------------------------------------------
# 2. Plot with fully adjusted aesthetics
# --------------------------------------------

ggplot(waffle_grid, aes(x = group_id, y = square_y)) +
  geom_tile(aes(fill = outcome), width = 0.8, height = 0.8, color = "black", linewidth = 0.3) +
  scale_fill_manual(values = c("Healed" = "black", "Not Healed" = "white")) +
  labs(
    title = "Wound Closure Outcomes by Treatment",
    subtitle = "Each square = 1 coral fragment",
    x = NULL, y = NULL, fill = "Wound Closure"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  scale_x_discrete(labels = c("Fish - Large", "No Fish - Large", "Fish - Small", "No Fish - Small")) +
  coord_cartesian(clip = "off") +
  theme_bw(base_size = 14) +
  theme(
    axis.text.x = element_text(size = 14, vjust = 2),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    legend.position = "right",
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    plot.title = element_text(size = 18, face = "bold"),
    plot.subtitle = element_text(size = 15),
    plot.margin = margin(10, 10, 10, 10)
  )




# 1. Summarize counts
waffle_data <- wound_closure_clean %>%
  mutate(outcome = ifelse(wound_close == "Yes", "Healed", "Not Healed"),
         # reorder fish so No Fish is left, Fish is right
         fish    = factor(fish, levels = c("No Fish", "Fish"))
  ) %>%
  count(fish, wound, outcome)

# 2. Compute per‐group percent healed
pct <- waffle_data %>%
  group_by(fish, wound) %>%
  summarise(
    total    = sum(n),
    healed   = sum(if_else(outcome == "Healed", n, 0)),
    pct_heal = healed / total * 100,
    .groups  = "drop"
  )

# 3. Plot with facets: No Fish on left, Fish on right
library(waffle)    
library(ggplot2)
library(dplyr)

# 1. Summarize counts
waffle_data <- wound_closure_clean %>%
  mutate(outcome = ifelse(wound_close == "Yes", "Healed", "Not Healed"),
         fish    = factor(fish, levels = c("No Fish", "Fish"))
  ) %>%
  count(fish, wound, outcome)

# 2. Compute per‐group percent healed for annotation
pct <- waffle_data %>%
  group_by(fish, wound) %>%
  summarise(
    total    = sum(n),
    healed   = sum(if_else(outcome == "Healed", n, 0)),
    pct_heal = healed / total * 100,
    .groups  = "drop"
  )

# 3. Plot
ggplot(waffle_data, aes(values = n, fill = outcome)) +
  geom_waffle(
    n_rows = 6,
    size   = 0.3,       # thickness of the border
    colour = "black"    # black outline around every square
  ) +
  facet_grid(wound ~ fish) +
  scale_fill_manual(
    values = c(
      "Healed"     = "#2b4b52",
      "Not Healed" = "white"    # white fill = “empty” box
    ),
    guide = guide_legend(title = "Outcome")
  ) +
  geom_text(
    data = pct,
    aes(x = Inf, y = Inf, label = paste0(round(pct_heal), "%")),
    hjust = 1.1, vjust = 1.5,
    size = 5, fontface = "bold",
    inherit.aes = FALSE
  ) +
  labs(
    title    = "Wound Closure Outcomes by Treatment",
    subtitle = "Each square = one coral fragment"
  ) +
  theme_void(base_size = 14) +
  theme(
    strip.text      = element_text(face = "bold", size = 12),
    legend.position = "bottom",
    plot.title      = element_text(face = "bold", size = 16),
    plot.subtitle   = element_text(size = 14)
  )


# --------------------------------------------
# 3. Save with optimal figure size
# --------------------------------------------

ggsave(here("figures","wound_closure", "wound_closure_waffle_manuscript_final.pdf"),
       width = 6, height = 6, bg="white",dpi=150)


```
# Models for healed/not healed

## Creating lmer models
```{r}
mod_int <- glmer(wound_binary ~ fish * wound + (1 | tank), family = binomial, data = wound_closure_glm)
# anova(fish_glm, type="II", test.statistic="F")


mod_add <- glmer(wound_binary ~ fish + wound + (1|tank) , family = binomial, data = wound_closure_glm)
# anova(wound_glm, test = "Chisq")

mod_fish <- glmer(wound_binary ~ fish + (1| tank), family = binomial, data = wound_closure_glm )
# anova(fish_glmer, type="II", test.statistic="F")

mod_wound <- glmer(wound_binary ~ wound + (1|tank), family = binomial, data = wound_closure_glm)

mod_null <- glmer(wound_binary ~ 1 + (1|tank), family = binomial, data = wound_closure_glm)

```

## Likelihood tests
```{r}
lrt_interaction <- anova(mod_add, mod_int, test = "Chisq")
lrt_fish <- anova(mod_wound, mod_add, test = "Chisq")
lrt_wound <- anova(mod_fish, mod_add, test = "Chisq")
```

## Likelihood table
```{r}
lrt_table <- tibble(
  Test = c("Interaction (Fish × Wound)", "Fish Effect", "Wound Effect"),
  ChiSq = c(
    lrt_interaction$Chisq[2],
    lrt_fish$Chisq[2],
    lrt_wound$Chisq[2]
  ),
  Df = c(
    lrt_interaction$Df[2],
    lrt_fish$Df[2],
    lrt_wound$Df[2]
  ),
  p_value = c(
    lrt_interaction$`Pr(>Chisq)`[2],
    lrt_fish$`Pr(>Chisq)`[2],
    lrt_wound$`Pr(>Chisq)`[2]
  )
) %>%
  mutate(across(where(is.numeric), round, 3))

# Format with gt
lrt_table %>%
  gt() %>%
  tab_header(
    title = "Table #. Likelihood Ratio Tests for Effects on wound healing",
    #subtitle = paste("Scaled growth:", expression((final - initial)/initial^b))
  ) %>%
  cols_label(
    Test = "Model Comparison",
    ChiSq = "Chi-squared",
    Df = "Degrees of Freedom",
    p_value = "P-value"
  ) %>%
  fmt_number(columns = vars(ChiSq, p_value), decimals = 3) %>%
  tab_options(
    table.font.size = 14,
    heading.title.font.size = 16,
    heading.subtitle.font.size = 13
  )
```

# Models for final proportion of wound remaining
#simultaneous model comparisons - best model?
```{r}
mod_full <- glmmTMB(
  prop_wound_remaining ~ fish * wound + (1 | tank),
  ziformula = ~ fish * wound,
  data = percent_closure,
  family = beta_family(link = "logit")
)


mod_no_interaction <- glmmTMB(
  prop_wound_remaining ~ fish + wound + (1 | tank),  # 
  ziformula = ~ fish + wound ,                         # 
  data = percent_closure,
  family = beta_family(link = "logit")
)


mod_no_fish <- glmmTMB(
  prop_wound_remaining ~ wound + (1 | tank),  # 
  ziformula = ~ wound ,                         # 
  data = percent_closure,
  family = beta_family(link = "logit")
)

mod_no_wound <- glmmTMB(
  prop_wound_remaining ~ fish + (1 | tank),  
  ziformula = ~ fish ,                         # 
  data = percent_closure,
  family = beta_family(link = "logit")
)


lrt_interaction_zero <- anova(mod_no_interaction, mod_full, test = "Chisq")
lrt_fish_zero <- anova(mod_no_fish, mod_no_interaction, test = "Chisq")
lrt_wound_zero <- anova(mod_no_wound, mod_no_interaction, test = "Chisq")

lrt_table <- tibble(
  Test = c("Interaction (Fish × Wound)", "Fish Effect", "Wound Effect"),
  ChiSq = c(lrt_interaction_zero$Chisq[2],
            lrt_fish_zero$Chisq[2],
            lrt_wound_zero$Chisq[2]),
  Df = c(lrt_interaction_zero$Df[2],
         lrt_fish_zero$Df[2],
         lrt_wound_zero$Df[2]),
  p_value = c(lrt_interaction_zero$`Pr(>Chisq)`[2],
              lrt_fish_zero$`Pr(>Chisq)`[2],
              lrt_wound_zero$`Pr(>Chisq)`[2])
) %>%
  mutate(across(where(is.numeric), round, 3))

lrt_table %>%
  gt() %>%
  tab_header(
    title = "Table #. Likelihood Ratio Tests for Effects on wound healing",
    #subtitle = paste("Scaled growth:", expression((final - initial)/initial^b))
  ) %>%
  cols_label(
    Test = "Model Comparison",
    ChiSq = "Chi-squared",
    Df = "Degrees of Freedom",
    p_value = "P-value"
  ) %>%
  fmt_number(columns = vars(ChiSq, p_value), decimals = 3) %>%
  tab_options(
    table.font.size = 14,
    heading.title.font.size = 16,
    heading.subtitle.font.size = 13
  )

```

# Models for healing rate
## Creating the Models
```{r}
rate_full <- lmer(
  healing_rate ~ fish * wound + (1 | tank),
  data = percent_closure
)


rate_no_interaction <- lmer(
  healing_rate ~ fish + wound + (1 | tank),   
  data = percent_closure
)


rate_no_fish <- lmer(
  healing_rate ~ wound + (1 | tank),   
  data = percent_closure
)

rate_no_wound <- lmer(
  healing_rate ~ fish + (1 | tank),  
  data = percent_closure
)


lrt_interaction_zero_rate <- anova(rate_no_interaction, rate_full, test = "Chisq")
lrt_fish_zero_rate <- anova(rate_no_fish, rate_no_interaction, test = "Chisq")
lrt_wound_zero_rate <- anova(rate_no_wound, rate_no_interaction, test = "Chisq")

lrt_table <- tibble(
  Test = c("Interaction (Fish × Wound)", "Fish Effect", "Wound Effect"),
  ChiSq = c(lrt_interaction_zero_rate$Chisq[2],
            lrt_fish_zero_rate$Chisq[2],
            lrt_wound_zero_rate$Chisq[2]),
  Df = c(lrt_interaction_zero_rate$Df[2],
         lrt_fish_zero_rate$Df[2],
         lrt_wound_zero_rate$Df[2]),
  p_value = c(lrt_interaction_zero_rate$`Pr(>Chisq)`[2],
              lrt_fish_zero_rate$`Pr(>Chisq)`[2],
              lrt_wound_zero_rate$`Pr(>Chisq)`[2])
) %>%
  mutate(across(where(is.numeric), round, 3))

lrt_table %>%
  gt() %>%
  tab_header(
    title = "Table #. Likelihood Ratio Tests for Effects on Healing Rate",
  ) %>%
  cols_label(
    Test = "Model Comparison",
    ChiSq = "Chi-squared",
    Df = "Degrees of Freedom",
    p_value = "P-value"
  ) %>%
  fmt_number(columns = vars(ChiSq, p_value), decimals = 3) %>%
  tab_options(
    table.font.size = 14,
    heading.title.font.size = 16,
    heading.subtitle.font.size = 13
  )
```

## R2 values

```{r}
r2_full <- r.squaredGLMM(rate_full)

r2_full
```


# Rough code

### figuring which model to use for final wound healing
```{r}
#modeling proportion of the wound remaining at Tf, this gets around the 1 inflation, and allows for 0 inflation, which glmmTMB can deal with. 
percent_closure$prop_wound_remaining <- 1 - percent_closure$prop
                              
beta_mod <- glmmTMB(
  prop_wound_remaining ~ fish * wound + (1 | tank),
  ziformula = ~ fish * wound,
  data = percent_closure,
  family = beta_family(link = "logit")
)

summary(beta_mod)

beta_mod_null<- glmmTMB(
  prop_wound_remaining ~ fish * wound + (1 | tank),
  ziformula = ~1,
  data = percent_closure,
  family = beta_family(link = "logit")
)

beta_mod_fix<- glmmTMB(
  prop_wound_remaining ~ fish * wound + (1 | tank),
  ziformula = ~fish + wound,
  data = percent_closure,
  family = beta_family(link = "logit")
)

beta_mod_fish<- glmmTMB(
  prop_wound_remaining ~ fish * wound + (1 | tank),
  ziformula = ~fish,
  data = percent_closure,
  family = beta_family(link = "logit")
)
summary(beta_mod)
AIC(beta_mod_null, beta_mod)
AIC(beta_mod_fish, beta_mod_fix)

summary(beta_mod_fish)
summary(beta_mod_null)




#GAMLSS analysis...I don't trust any of these models. Partly because I proabably don't understand and because they're not giving me the values I want :(
beta_prop_1 <- gamlss(
  prop ~ fish * wound + random(tank),    
  sigma.fo = ~1,                        
  nu.fo    = ~1,                        
  tau.fo   = ~ fish * wound,           
  family   = BEINF,
  data     = percent_closure
)


mod <- gamlss(
  prop ~ fish * wound + random(tank),
  family = BEINF,
  data = percent_closure
)

beta_prop_null <- gamlss(
  prop ~ fish * wound + random(tank),    
  sigma.fo = ~1,                        
  nu.fo    = ~1,                        
  tau.fo   = ~1,           
  family   = BEINF,
  data     = percent_closure
)

summary(beta_prop_null)
summary(mod)

AIC(beta_prop_1, beta_prop_null)
anova(beta_prop_1, beta_prop_null)

```

### zero - model comparisons
```{r}
mod_full_zero <- glmmTMB(
  prop_wound_remaining ~ fish * wound + (1 | tank),
  ziformula = ~ fish * wound,
  data = percent_closure,
  family = beta_family(link = "logit")
)

summary(mod_full)

mod_no_interaction_zero <- glmmTMB(
  prop_wound_remaining ~ fish * wound + (1 | tank),  
  ziformula = ~ fish + wound,                         
  data = percent_closure,
  family = beta_family(link = "logit")
)

mod_no_fish_zero <- glmmTMB(
  prop_wound_remaining ~ fish * wound + (1 | tank),  
  ziformula = ~ wound,                      
  data = percent_closure,
  family = beta_family(link = "logit")
)

mod_no_wound_zero <- glmmTMB(
  prop_wound_remaining ~ fish * wound + (1 | tank),  
  ziformula = ~ fish,                         
  data = percent_closure,
  family = beta_family(link = "logit")
)

lrt_interaction <- anova(mod_no_interaction_zero, mod_full_zero, test = "Chisq")
lrt_fish <- anova(mod_no_fish_zero, mod_no_interaction_zero, test = "Chisq")
lrt_wound <- anova(mod_no_wound_zero, mod_no_interaction_zero, test = "Chisq")

lrt_table <- tibble(
  Test = c("Interaction (Fish × Wound)", "Fish Effect", "Wound Effect"),
  ChiSq = c(lrt_interaction$Chisq[2],
            lrt_fish$Chisq[2],
            lrt_wound$Chisq[2]),
  Df = c(lrt_interaction$Df[2],
         lrt_fish$Df[2],
         lrt_wound$Df[2]),
  p_value = c(lrt_interaction$`Pr(>Chisq)`[2],
              lrt_fish$`Pr(>Chisq)`[2],
              lrt_wound$`Pr(>Chisq)`[2])
) %>%
  mutate(across(where(is.numeric), round, 3))

lrt_table
```


### model comparisons with zero kept to fixed effects
```{r}
mod_full <- glmmTMB(
  prop_wound_remaining ~ fish * wound + (1 | tank),
  ziformula = ~ fish + wound,
  data = percent_closure,
  family = beta_family(link = "logit")
)

summary(mod_full)

mod_no_interaction <- glmmTMB(
  prop_wound_remaining ~ fish + wound + (1 | tank),  
  ziformula = ~ fish + wound,                        
  data = percent_closure,
  family = beta_family(link = "logit")
)

# Remove fish effect
mod_no_fish <- glmmTMB(
  prop_wound_remaining ~ wound + (1 | tank),  
  ziformula = ~ fish + wound,                         
  data = percent_closure,
  family = beta_family(link = "logit")
)

# Remove wound effect
mod_no_wound <- glmmTMB(
  prop_wound_remaining ~ fish + (1 | tank),  
  ziformula = ~ fish + wound,                         
  data = percent_closure,
  family = beta_family(link = "logit")
)

lrt_interaction <- anova(mod_no_interaction, mod_full, test = "Chisq")
lrt_fish <- anova(mod_no_fish, mod_no_interaction, test = "Chisq")
lrt_wound <- anova(mod_no_wound, mod_no_interaction, test = "Chisq")

lrt_table <- tibble(
  Test = c("Interaction (Fish × Wound)", "Fish Effect", "Wound Effect"),
  ChiSq = c(lrt_interaction$Chisq[2],
            lrt_fish$Chisq[2],
            lrt_wound$Chisq[2]),
  Df = c(lrt_interaction$Df[2],
         lrt_fish$Df[2],
         lrt_wound$Df[2]),
  p_value = c(lrt_interaction$`Pr(>Chisq)`[2],
              lrt_fish$`Pr(>Chisq)`[2],
              lrt_wound$`Pr(>Chisq)`[2])
) %>%
  mutate(across(where(is.numeric), round, 3))

lrt_table
```

